Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2018-05-01 LaTeX/response.tex

\noindent\paragraph{Response}
%\textcolor{red}{The abstract and summary still seem to be a little over-selling...}
\textcolor{blue}{
We have performed an extensive re-arrangement to improve the clarity of our writing, especially regarding the main results. 
As described in Response 2.1 (point 6) and below, we now provide a prominent and longer discussion of Kozielska et al 2010 and Ubeda et al 2015 (L57-69) and more precisely indicate which of our model features and results are new and unexpected in light of these previous results (see Response 2.2).
While related, these two studies considered particular cases that we do not consider at length. 
For example, Ubeda et al consider ESD-GSD transitions. 
Nevertheless, these transitions involve increased linkage with a haploid selected locus, which is precedent for our Conclusion 2A (which considers only GSD-GSD transitions).
We now mention this in the Introduction (L59-61) and again in the Discussion (L432).  
Similarly, Kozielska et al involves a special case of GSD transitions during which sex-linkage goes from perfect to none, which is is an extreme example of our Conclusion 2B.
This is twice mentioned (L64-67, L237). 
Despite this precedent, we find the permissive parameter space allowing GSD-GSD transitions surprising. 
%Especially because their main conclusion involved the strong influence of sex-ratio selection, which is explicitly not a driver during many such transitions. 
In addition, our Conclusions 3 and 4 reconcile Kozielska et al and Ubeda et al, explaining that sex-ratio selection is not an overwhelming force driving transitions in sex determination. 
We believe that our new presentation makes the novel and unexpected features of our results much more apparent. 
%To be clear, we do not claim to be the first to include both haploid and diploid selection to a model investigating the evolution of new sex-chromosomes and we do not consider the addition of sex-specific selection or analytical results for previously numerical models to be our most significant contributions (again, see Responses 2.1 and 2.2 and below for more detail on what these are). 
}

@@

\noindent\subsubsection{}
The findings in [SOO] claimed to be surprising (``Surprisingly, we find that neither force (sex ratio selection nor associations with genes that have sex-specific effects) dominates the spread of new sex-determining systems alone.") ... [but this result is rather similar to those in Ubeda et al 2015].

\noindent\paragraph{Response}
\textcolor{blue}{
Here we were referring to the result that $\lambda_{W'}^{XY} = \lambda_{Y'}^{ZW}$ with haploid selection, even when sex-ratio biases evolve to become either stronger or weaker (e.g., $\lambda_{Y'}^{ZW}$ vs. $\lambda_{W'}^{XY}$ with $R<r$ and male meiotic drive) or when sex-ratio biases are either present or absent (e.g., $\lambda_{W'}^{XY}$ vs. $\lambda_{Y'}^{ZW}$ with $r<1/2$, $R=1/2$, and male meiotic drive).  
Step 2 in Ubeda et al 2015 shows that associations with a haploid selected locus can favour a new sex-determining allele despite causing biased sex ratios. 
Nevertheless, it was difficult for us to intuit whether the presence/absence of ancestral sex-ratio bias would have a negligible or overwhelming effect on the evolution of new sex-determining alleles. 
For example, step 3 of Ubeda et al involves the spread of an autosomal meiotic drive suppressor favoured by Fisherian sex-ratio selection and Kozielska et al 2010 show that ancestral sex-ratio bias can favour new sex determining systems that balance the sex ratio.
These previous results are now discussed in the Introduction (L57-69). 
We found our result, that these two forces are exactly equal (with selection weak relative to recombination), such that there is no difference between the propensity for XY-ZW or ZW-XY transitions, to be surprising.
To clarify our meaning we have rephrased (in Author summary): ``Surprisingly, we find the two forces (selection to equalize the sex-ratio and the benefits of hitch-hiking with driven alleles that distort the sex ratio) will often be equally strong, and thus neither is sufficient to explain the spread of new sex-determining systems in every case." 
We believe that the re-framing of the Introduction/Discussion and the emphasis given to specific Conclusions 1-4 will help to clarify this point throughout the paper.
}

@@

\noindent\subsubsection{}
\noindent and unexpected (``Even more unexpectedly, we find that, to spread, new sex-determining alleles do not necessarily have to arise in closer linkage with genes that are differentially selected in males and females"), are rather similar to those in Ubeda et al 2015. This is a problem that repeats itself over the paper

\noindent\paragraph{Response}
\textcolor{blue}{
The Ubeda et al 2015 results do indeed show that differential selection between male and female diploids is not required for new sex-determining alleles to spread; they show that a new genetic sex-determining allele can spread from ancestral ESD when it is linked to a locus that experiences sex-specific meiotic drive. 
In this sentence, we were attempting to stress that new sex determining alleles can spread when they are \textit{less closely linked or even completely unlinked} to loci under differential male/female haploid and/or diploid selection. 
To our knowledge, this has only been shown for the special case of male meiotic drive in an ancestral XY system with $r=0$ and $R=1/2$ with a neo-W, which is driven by sex-ratio selection (Kozielska et al 2010). 
We therefore found the large parameter space allowing invasion with $r<R$, sometimes in the absence of sex-ratio selection (or even in the absence of haploid selection), to be very unexpected.
Nevertheless, we have rephrased (in Author Summary): ``We also find that new sex-determining alleles can spread despite being less closely linked to selected loci, driven by initially tight linkage or haploid selection." 
The changes to the presentation of the Introduction and Results should also help to clarify this point further (see Responses 2.1 and 2.2).
%We discuss the related results of Kozielska and Ubeda on lines 261 and 350.
}

@@

\noindent\subsubsection{}
For example in: Page 2. The authors claim that predominant theories of the evolution of sex-determining systems are two: sexually-antagonistic selection and sex-ratio selection. They ignore other two: inbreeding (Charlesworth 1978) and sex-different drive (Ubeda et al 2015).

\noindent\paragraph{Response}
\textcolor{blue}{
First, we believe sex-differences in selection and sex ratio selection are two ``predominant" theories explaining sex-determination transitions. 
Second, sexually-antagonistic selection was referred to as an example of ``selective differences between the sexes". 
We believe the phrase 'selective differences between the sexes' includes sex-differences in drive (Ubeda et al 2015), which just occurs during haploid rather than diploid selection (see also Response 3.2.6).
Finally, Charlesworth and Charlesworth 1978 (Am Nat and Heredity) discuss the role of inbreeding (and particularly inbreeding depression) in transitions from hermaphroditism to dioecy, without regard for genetic sex-determination. 
As noted on lines 468-472 in the original submission, we have only considered GSD-GSD transitions and GSD-ESD transitions.
For clarity, we have moved this description into the Introduction (L49-50). 
In addition, where we draw a parallel between ESD and hermaphroditism (L468-476), we suggest future theory include inbreeding, haploid selection, and sex chromosomes. 
}

@@

%\textcolor{blue}{
%We are unsure of which paper the reviewer is referring to.
%Charlesworth 1978 PNAS 75:5618 discusses the degradation of the Y by Muller's ratchet and subsequent dosage compensation, which has little relevance here.
%Charlesworth and Charlesworth 1978 Am Nat 112:975-997 discuss the role of inbreeding in transitions from hermaphrodism (ie no sex-determining alleles) to the case of dioecy (with sex-determining alleles). 
%Charlesworth and Charlesworth 1978 Heredity 41:137-153 is similar but begins with monoecy.
%Assuming the reviewer meant one of the latter two, both it and Ubeda et al 2015 consider the origin of genetic sex-determination from hermaphrodism/monoecy or environmental sex-determination. 
%Here we are much more concerned with transitions between genetic sex-determination systems.
%Further, we do not state that sex-antagonism is one of the predominant theories explaining SD transitions, it is but one example where fitness differs between the sexes, another example being ``sex-different drive". 
%We discuss the fact that sex-specific drive can be thought of as a sex difference in selection in paragraph 5 of the introduction (see Response 3.2.6).
%We have rephrased slightly to restrict ourselves to transitions between GSD: ``Predominant theories accounting for transitions between GSD systems by selection involve fitness differences between sexes (e.g., sexually antagonistic selection) or sex-ratio selection...".}

\noindent\subsubsection{}
Page 3 ``Here we use mathematical models to find the conditions under which new sex-determining systems spread when individuals experience selection at both diploid and haploid stages, which allows fitness differences between the sexes and sex ratio biases to occur simultaneously".
This would be the same model as Ubeda et al 2015 adding sex-specific viability.

\noindent\paragraph{Response}
\textcolor{blue}{
As we now make clear, we do not look at the same situation as Ubeda et al 2015 and our contribution is not just to add sex-specific viability. 
In particular, Ubeda et al start from ESD. 
Therefore, associations with an ancestral sex-determining locus cannot build up. 
These lead to some of our most surprising results (e.g., haploid selection favours trans-GSD transitions for a wider variety of genomic locations than diploid selection alone; ESD does not invade despite initial sex-ratio bias).
Further, our model and intuition-giving analytical solutions explore a much wider range of scenarios than the numerical results of Ubeda et al, allowing, for example, sex-specific selection (e.g., sex-antagonism) and various genomic arrangements (any order/linkage of the three loci).
Our model is thus much more general and clarifies just when one should expect to see the numerical results of Ubeda et al.
We have rephrased to make this clear (after discussing Ubeda et al 2015 in the Introduction, see Response 2.1; L70-74): ``Here, we analytically find the conditions under which new GSD or ESD systems spread in ancestral GSD systems with generic linkage between the loci involved and generic sex-specific haploid and diploid selection. 
Doing so, we reconcile the results of Kozielska et al.\ (2010) and Ubeda et al.\ (2015) by showing when new GSD systems that increase \textit{or} decrease linkage with loci under haploid selection spread."
}

@@

\noindent\subsubsection{}
Lines 68-69: ``If we assume that haploid selection at any particular locus predominantly occurs in one sex (e.g., meiotic drive during spermatogenesis), then such loci experience a form of sex-specific selection."

This is a stretch and creates confusion. I consider selection acting on individuals with different sex-specific selection, now when selection is acting on gametes (which are neither males or females) males or females are just different environments in which selection may take place. Neither the dynamics nor the constraints are the same.

\noindent\paragraph{Response}
\textcolor{blue}{In this study, we are referring to organisms, such as animals and angiosperms, in which the type of gamete depends on the sex of the diploid that produced it. In this respect, one could consider gametic selection among sperm, for example, to be a form of male-specific selection. However, as noted here, the dynamics of selection during gametic selection cannot be captured by a diploid selection coefficient, which is why we model selection upon haploid genotypes specifically in two ways. }
\textcolor{blue}{
We have attempted to clarify by expanding our explanation of the mating system and the way we model haploid selection (see Response 3.2 below).}

@@

\noindent\subsubsection{}
Lines 70-71: ``In this respect, we might expect that haploid selection would affect transitions between sex-determining systems in a similar manner to sex-specific diploid selection (as explored in [31,32])"

Following up my previous comment Ubeda 2015 explores this and shows that both forms of selection are different. For example sex-antagonistic selection is needed for the invasion of a new sex-determining gene with diploid selection while antagonistic drive in each of the sexes is not needed for the invasion of a new sex-determining gene with haploid selection.

\noindent\paragraph{Response}
\textcolor{blue}{In the absence of sex-differences in selection it is true that antagonistic drive is not needed, but \textit{differences} in haploid selection between the sexes are needed (e.g., drive in only one sex, maintained by ploidally-antagonistic selection). 
If drive was the same in both sexes, $\alpha^{m}=\alpha^{f}$ (and all other male/female selection terms also equal), we show that the new sex-determining allele would not invade (e.g., equation 2 and 3). 
In this sense haploid selection is similar to sex-specific diploid selection, and together haploid and diploid selection determine the differences in male vs.\ female fitness, which can drive sex-determination transitions (see Response 3.2.4).
This sentence meant to introduce the `sex-specific' aspect of haploid selection in the way Ubeda et al 2015 considers it; associations with alleles favoured in one sex (or the gametes produced by that sex) should favour a new sex-determining allele.
We have rephrased to make this clear (L52-54): ``Therefore, one might expect new sex-determining systems to benefit from close linkage with haploid selected loci, as found for sex-differences in diploid selection."
We now follow this sentence up with a paragraph that clarifies its meaning by comparing with the results of Kozielska et al 2010 and Ubeda et al 2015 (see Response 2.1).}

@@

\noindent\subsubsection{}
Lines 77-78: ``It is not immediately clear how the spread of new sex-determining systems would be influenced by the combination of sex ratio biases and associations with haploid selected allele"

This is indeed the focus of Ubeda et al 2015.

\noindent\paragraph{Response}
\textcolor{blue}{While Ubeda et al 2015 did focus on sex ratio biases and haploid selected alleles, the system started with ESD, where there are no associations (see Response 3.1.4).
Further, the restricted scope of that numerical study prevents us from generalizing beyond its special case (see Response 3.1.4).
We now present a refurbished argument for why it is difficult to intuit the implications of haploid selection while summarizing the results of Ubeda et al and Kozielska et al 2010 in the Introduction (see Response 2.1).
}

@@

\noindent\subsubsection{}
I will not provide further examples, but it is surprising thatI the work of Ubeda et al 2015, which is closely related to the one presented here, is cited in page 12 supporting a statement only tangential to the work of Ubeda et al 2015.

\noindent\paragraph{Response}
\textcolor{blue}{Ubeda et al 2015 was already cited on line 396 of the original submission, where we mentioned that it showed that sex-ratio biases could become more severe during a transition.
Nevertheless, we now discuss Ubeda et al 2015 in the Introduction, Results, and Discussion (see Responses above).}

@@

\noindent\subsection{}
In terms of modeling the distinction between meiotic drive and gametic competition is unnecessary and only complicates notation. Furthermore, from biological perspective the division seems blurry with molecular biologists considering that meiotic drive only happens in females and never in males and what is often called meiotic drive in males is gamete competition. I would suggest using a single parameter measuring transmission distortion.

\noindent\paragraph{Response}
\textcolor{blue}{We do not agree. 
Given our definition of meiotic drive and gametic competition (L106-116), the two processes are different.
In particular, meiotic drive is assumed to occur during meiosis, and thus is ``gametic competition" amongst the gametes produced by a single individual. 
As such, meiotic drive can only have an effect in heterozygotes.
Gametic competition on the other hand, is competition amongst all the gametes produced by one sex (e.g., pollen from many male plants).
Thus allele frequency changes can occur by gametic competition even if all individuals were homozygous (as long as there is a polymorphism).
We highlight these differences in the model section (L106-116).
}
